Robust Co-occurrence Quantification for Lexical Distributional Semantics

### 2016-06-10 paper.tex
Updates from ShareLaTeX

% \aclfinalcopy % Uncomment this line for the final submission
\def\aclpaperid{30} %  Enter the acl Paper ID here

@@

%\renewcommand{\baselinestretch}{0.95}

@@

Words that occur in similar context have similar meaning \cite{harris1954distributional}. Thus the meaning of a word can be modeled by counting its co-occurrence with neighboring words in a corpus. Distributional models of meaning represent co-occurrence information in a vector space, where the dimensions are the neighboring words and the values are co-occurrence counts. Successful models need to be able to discriminate co-occurrence information, as not all co-occurrence counts are equally useful, for instance, the co-occurrence with the article \textit{the} is less informative than with the noun \textit{existence}. Another fundamental question is the vector space dimensionality and what neighbor words should correspond to them.

@@

An alternative way of obtaining a vector space with few dimensions, usually with just 100--500, is the use of SVD as a part of Latent Semantic Analysis \cite{ARIS:ARIS1440380105} or another models such as SGNS \cite{mikolov2013efficient} and GloVe \cite{pennington2014glove}. However, these models take more time to instantiate in comparison to weighting of a co-occurrence matrix, bring more parameters to explore and produce vector spaces with uninterpretable dimensions. In this work we focus on vector spaces that directly weight a co-occurrence matrix and report results for SVD, GloVe and SGNS from the study of \newcite{TACL570} for comparison.

The mismatch of experiments with non-dense models in vector dimensionality between lexical and compositional tasks gives rise to a number of questions:
\begin{compactitem}
\item To what extent does model performance depend on vector dimensionality?
\item Do parameters influence 200K and 1K dimensional models similarly? Can the findings of \newcite{TACL570} be directly applied to models with a few thousand dimensions?
\item If not, can we derive suitable parameter selection heuristics which take account of dimensionality?
\end{compactitem}

To answer these questions, we perform a systematic study of distributional models with a rich set of parameters on SimLex-999  \cite{hill2014simlex}, a lexical \emph{similairty} dataset, and test selected models on MEN \cite{Bruni:2014:MDS:2655713.2655714}, a lexical \emph{relatedness} dataset. Both datasets are currently widely used and surpass datasets stemming from information retrieval, WordSim-353 \cite{2002:PSC:503104.503110}, and computational linguistics, RG65 \cite{Rubenstein:1965:CCS:365628.365657}, in quantity by having more entries and quality by attention to evaluated relation.

@@

Many approaches use only \emph{positive} PMI values, as  negative PMI values may not positively contribute to model performance and sparser matrices are more computationally tractable \cite{Turney:2010:FMV:1861751.1861756}. This can be generalised to an additional cutoff parameter $k$ (\texttt{neg}) following \newcite{TACL570}, giving our third PMI variant (abbreviated as SPMI):
%
\begin{equation}
  \label{eq:ppmi}
  \operatorname{SPMI_k} = \max (0, \operatorname{PMI}(x, y) - \log k)
\end{equation}
%
When $k = 1$ SPMI is equivalent to positive PMI. $k > 1$ increases the underlying matrix sparsity by keeping only highly associated co-occurrence pairs. $k < 1$ decreases the underlying matrix sparsity by including some unassociated co-occurrence pairs, which are usually excluded due to unreliability of probability estimation \cite{Dagan:1993:CWS:981574.981596}.

@@

\section{Experiments}
\label{sec:lexical-experiments}

@@

@@

\paragraph{Evaluation of heuristics}
\label{sec:heuristic-evaluation}

We evaluate these heuristics by comparing the performance they give on SimLex-999 against that obtained using the best possible parameter selections (determined via an exhaustive search at each dimensionality setting). We also compare them to the best scores reported by \newcite{TACL570} for their model (PMI and SVD), word2vec-SGNS \cite{mikolov2013efficient} and GloVe \cite{pennington2014glove} -- see Figure~\ref{fig:best-simlex} (only the better-performing SPMI and SCPMI are shown).

@@

Overall, $\log n$ weighting should be used with PMI, CPMI and SCPMI. High-dimensional SPMI models show the same behaviour, but if $D<$ 10K, no weighting should be applied. SPMI and SCPMI should be preferred over CPMI and PMI. As Figure~\ref{fig:global-best-simlex} shows, our heuristics give performance close to the optimum for any dimensionality, with a large improvement over both an average parameter setting and the parameters suggested by \cite{TACL570} in a high-dimensional setting.\footnote{Our results using \newcite{TACL570}'s parameters differ slightly from theirs due to different window sizes (5 vs 2).}

@@

@@

We thank the anonymous reviewers for their comments. Support from EPSRC grant EP/J002607/1 is gratefully acknowledged by Milajevs and Sadrzadeh.
Purver is partly supported by ConCreTe: the project ConCreTe acknowledges the financial support of the Future and Emerging Technologies (FET) programme within the Seventh Framework Programme for Research of the European Commission, under FET grant number 611733.

>>>>

\aclfinalcopy % Uncomment this line for the final submission
\def\aclpaperid{30} %  Enter the acl Paper ID here

@@

% \renewcommand{\baselinestretch}{0.95}

@@

Words that occur in similar context have similar meaning \cite{harris1954distributional}. Thus the meaning of a word can be modeled by counting its co-occurrence with neighboring words in a corpus. Distributional models of meaning represent co-occurrence information in a vector space, where the dimensions are the neighboring words and the values are co-occurrence counts. Successful models need to be able to discriminate co-occurrence information, as not all co-occurrence counts are equally useful, for instance, the co-occurrence with the article \textit{the} is less informative than with the noun \textit{existence}. The discrimination is usually achieved by weighting of co-occurrence counts. Another fundamental question in vector space design is the vector space dimensionality and what neighbor words should correspond to them.

@@

An alternative way of obtaining a vector space with few dimensions, usually with just 100--500, is the use of SVD as a part of Latent Semantic Analysis \cite{ARIS:ARIS1440380105} or another models such as SGNS \cite{mikolov2013efficient} and GloVe \cite{pennington2014glove}. However, these models take more time to instantiate in comparison to weighting of a co-occurrence matrix, bring more parameters to explore and produce vector spaces with uninterpretable dimensions (vector space dimension interpretation is used by some lexical models, for example, \newcite{McGregorEtAl15JAGI}, and the passage from formal semantics to tensor models relies on it \cite{DBLP:journals/corr/abs-1003-4394}). In this work we focus on vector spaces that directly weight a co-occurrence matrix and report results for SVD, GloVe and SGNS from the study of \newcite{TACL570} for comparison.

The mismatch of recent experiments with non-dense models in vector dimensionality between lexical and compositional tasks gives rise to a number of questions:
\begin{compactitem}
\item To what extent does model performance depend on vector dimensionality?
\item Do parameters influence 200K and 1K dimensional models similarly? Can the findings of \newcite{TACL570} be directly applied to models with a few thousand dimensions?
\item If not, can we derive suitable parameter selection heuristics which take account of dimensionality?
\end{compactitem}

To answer these questions, we perform a systematic study of distributional models with a rich set of parameters on SimLex-999  \cite{hill2014simlex}, a lexical \emph{similairty} dataset, and test selected models on MEN \cite{Bruni:2014:MDS:2655713.2655714}, a lexical \emph{relatedness} dataset. The reason for choosing these is that they are currently widely used and surpass datasets stemming from information retrieval, WordSim-353 \cite{2002:PSC:503104.503110}, and computational linguistics, RG65 \cite{Rubenstein:1965:CCS:365628.365657}, in quantity by having more entries and in quality by attention to evaluated relations.

@@

Many approaches use only \emph{positive} PMI values, as  negative PMI values may not positively contribute to model performance and sparser matrices are more computationally tractable \cite{Turney:2010:FMV:1861751.1861756}. This can be generalised to an additional cutoff parameter $k$ (\texttt{neg}) following \newcite{TACL570}, giving our third PMI variant (abbreviated as SPMI):\footnotemark{}
%
\begin{equation}
  \label{eq:ppmi}
  \operatorname{SPMI_k} = \max (0, \operatorname{PMI}(x, y) - \log k)
\end{equation}
%
When $k = 1$ SPMI is equivalent to positive PMI. $k > 1$ increases the underlying matrix sparsity by keeping only highly associated co-occurrence pairs. $k < 1$ decreases the underlying matrix sparsity by including some unassociated co-occurrence pairs, which are usually excluded due to unreliability of probability estimates \cite{Dagan:1993:CWS:981574.981596}.

\footnotetext{SPMI is different from CPMI because $\log\frac{P(x, y)}{P(x)P(y)} - \log{}k = \log\frac{P(x, y)}{P(x)(P(y)k} \neq \log\Big( 1 + \frac{P(x,y)}{P(x)P(y)} \Big)$.}

@@

\section{Experimental setup}
\label{sec:lexical-experiments}

@@

\section{Heuristics}

@@

\section{Evaluation of heuristics}
\label{sec:heuristic-evaluation}

We evaluate these heuristics by comparing the performance they give on SimLex-999 against that obtained using the best possible parameter selections (determined via an exhaustive search at each dimensionality setting). We also compare them to the best scores reported by \newcite{TACL570} for their model (PMI and SVD), word2vec-SGNS \cite{mikolov2013efficient} and GloVe \cite{pennington2014glove}---see Figure~\ref{fig:best-simlex}, where only the better-performing SPMI and SCPMI are shown.

@@

Overall, $\log n$ weighting should be used with PMI, CPMI and SCPMI. High-dimensional SPMI models show the same behaviour, but if $D<$ 10K, no weighting should be applied. SPMI and SCPMI should be preferred over CPMI and PMI. As Figure~\ref{fig:global-best-simlex} shows, our heuristics give performance close to the optimum for any dimensionality, with a large improvement over both an average parameter setting and the parameters suggested by \newcite{TACL570} in a high-dimensional setting.\footnote{Our results using \newcite{TACL570}'s parameters differ slightly from theirs due to different window sizes (5 vs 2).}

@@

There are several directions for the future work. Our experiments show that models with a few thousand dimensions are competitive with more dimensional models, see Figure~\ref{fig:best}. Moreover, for these models, unsmoothed probabilities give the best result. It might be the case that due to the large size of the corpus used, the probability estimates for the most frequent words are reliable without smoothing. More experiments need to be done to see whether this holds for smaller corpora.

The similarity datasets are transferred to other languages \cite{DBLP:journals/corr/LeviantR15}. The future work might investigate whether our results hold for languages other than English.

The qualitative influence of the parameters should be studied in depth with extensive error analysis on how parameter selection changes similarity judgements.

@@

We thank Ann Copestake for her valuable comments as part of the ACL SRW mentorship program and the anonymous reviewers for their comments. Support from EPSRC grant EP/J002607/1 is gratefully acknowledged by Dmitrijs Milajevs and Mehrnoosh Sadrzadeh.
Matthew Purver is partly supported by ConCreTe: the project ConCreTe acknowledges the financial support of the Future and Emerging Technologies (FET) programme within the Seventh Framework Programme for Research of the European Commission, under FET grant number 611733.
